---
title: "elt2_promoter_microscopy"
author: "Robert Williams"
date: "3/17/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install Packages
Install the necessary packages. Do this once the first time analysis is performed.   Uncomment and execute the following code chunk.

```{r}
# install.packages("tidyverse")
# install.packages("readxl")
# install.packages("ggpubr")
# install.packages("mratios")
```


## Load Package Libraries

```{r echo=FALSE}
library(tidyverse)
library(readxl)
library(ggpubr)
library(mratios)
```

# Load in quantification data

Strains:
- JM149 = elt-2 promoter reporter, caIs71[elt-2p::GFP::HIS-2B::unc-54 3'UTR + rol-6(su1006)]
- JM259 = elt-2 promoter reporter + elt-7 delection, elt-7(tm840) V; caIs71[elt-2p::GFP::HIS-2B::unc-54 3'UTR + rol-6(su1006)]

Each worm should have the following measurements:

- One intestine fluorescence measurement
- Four background measurements

The input data should have the following columns (without quotes):

 - "treatment": The type of RNAi treatment
 - "life_stage": The worm developmental stage
 - "worm": the worm number imaged
 - "measurement_type": `background` or `GFP`
 - "measurement_num": the measurement number from ImageJ
 - "area": region of interest area
 - "mean": mean gray pixel intensity
 - "min": minimum pixel intensity
 - "max": maximum pixel intensity
 - "intDen": product of area and sum of the values of the pixels in the selection
 - "rawIntDen": sum of the values of the pixels in the selection
 - "experiment": experiment name including rep number
 - "strain": strain or reporter gene name
 
More info here: https://imagej.nih.gov/ij/docs/guide/146-30.html


```{r}
sheets <- c("Rep1_ELT-2_RNAi",  "Rep1_L4440_RNAi", "Rep2_ELT-2_RNAi",  "Rep2_L4440_RNAi", "Rep3_ELT-2_RNAi",  "Rep3_L4440_RNAi")

image_df_JM259 <- data.frame()
for(sheet in sheets) {
 toappend <- read_excel("../01_input/JM259_Reps123_elt-2_Promoter_ImageJ_Analysis.xlsx", sheet = sheet)
 toappend <- toappend %>% mutate(experiment = sheet)
 toappend$worm <- as.character(toappend$worm)
 image_df_JM259 <- bind_rows(image_df_JM259, toappend)
}
image_df_JM259$strain <- "JM259"

image_df_JM149 <- data.frame()
for(sheet in sheets) {
 toappend <- read_excel("../01_input/JM149_Reps123_elt-2_Promoter_ImageJ_Analysis.xlsx", sheet = sheet)
 toappend <- toappend %>% mutate(experiment = sheet)
 toappend$worm <- as.character(toappend$worm)
 image_df_JM149 <- bind_rows(image_df_JM149, toappend)
}

image_df_JM149$strain <- "JM149"


image_df<- image_df_JM149 %>% bind_rows(image_df_JM259)
image_df <- image_df %>% filter(life_stage == "L1")
head(image_df)

colnames(image_df)
# Set the factor levels for ordering in downstream plotting
image_df$treatment <- factor(image_df$treatment, levels = c("L4440", "ELT-2"))
# Give each worm a unique ID
image_df <- image_df %>%
  mutate(ID = paste(strain, experiment, life_stage, worm, sep = "_"))
```


# Calculate corrected total cell fluorescence (CTCF)

CTCF = Integrated Density – (Area of selected cell X Mean fluorescence of background readings)  
More information here: https://theolb.readthedocs.io/en/latest/imaging/measuring-cell-fluorescence-using-imagej.html

```{r}
# Take the mean of the background measurements
background_df<- image_df %>%
  group_by(strain, treatment, worm, measurement_type, life_stage, experiment, ID) %>%
  summarize(Mean_Background = mean(mean)) %>% 
  filter(measurement_type == "background") %>%
  ungroup() %>%
  dplyr::select(ID, Mean_Background)
background_df
```
```{r}
# calculate CTCF for each worm
image_ctcf <- image_df %>%
  filter(measurement_type == "GFP") %>%
  dplyr::select(strain, ID, treatment, worm, area, life_stage, experiment, intDen) %>%
  inner_join(background_df, by = "ID") %>%
  rowwise() %>%
  mutate(CTCF = intDen - (area*Mean_Background) )
image_ctcf
```

# Plot the quantification results

```{r fig.width=5, fig.height=4}
ggplot(image_ctcf, aes(treatment, CTCF/area)) + 
  geom_boxplot(width = 0.5) +
  geom_jitter(width = 0.25, alpha = 0.5, shape = 16) +
  facet_grid(.~strain) +
  labs(
    title = "Overexpression of ELT-2\npromoter is ELT-7 dependent",
    x = "RNAi Treatment",
    y = "Corrected Total Cell \nFluorescence (A.U.)"
  ) +
  stat_compare_means(comparisons = list(c("L4440", "ELT-2")), method = "t.test") +
  scale_x_discrete(labels = c("L4440", "ELT-2"))+
  theme_classic() +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 1),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size=18),
    plot.title = element_text(hjust = 0.5, size = 18)
    ) 

ggsave("../03_output/ELT-2_promoter_analysis_plot_210317.pdf", width = 5, height = 4, useDingbats=FALSE)  
```

# Measure fold change 
To see how much brighter ELT-2 is compared to L4440

```{r}
mean_ctcf <-
  image_ctcf %>%
  group_by(strain, treatment) %>%
  summarise(avgCTCF = mean(CTCF), sdCTCF = sd(CTCF))

mean_ctcf <-
  mean_ctcf %>% pivot_wider(names_from = treatment,
                            values_from = c(avgCTCF, sdCTCF))
mean_ctcf
```
# Calculate fold change and add confidence intervals

```{r}
image_ctcf %>%filter(life_stage=='L1' & treatment == 'ELT-2', strain == "JM149") %>% dplyr::select(CTCF) -> JM149_numerator
image_ctcf %>% filter(life_stage=='L1' & treatment == 'L4440', strain == "JM149") %>% dplyr::select(CTCF) -> JM149_denominator
ttestratio(JM149_numerator[[1]], JM149_denominator[[1]]) -> L1_ttest_ratio_JM149

image_ctcf %>%filter(life_stage=='L1' & treatment == 'ELT-2', strain == "JM259") %>% dplyr::select(CTCF) -> JM259_numerator
image_ctcf %>% filter(life_stage=='L1' & treatment == 'L4440', strain == "JM259") %>% dplyr::select(CTCF) -> JM259_denominator
ttestratio(JM259_numerator[[1]], JM259_denominator[[1]]) -> L1_ttest_ratio_JM259
    
mean_ctcf$lower = c(L1_ttest_ratio_JM149$conf.int[1], L1_ttest_ratio_JM259$conf.int[1])
mean_ctcf$upper = c(L1_ttest_ratio_JM149$conf.int[2], L1_ttest_ratio_JM259$conf.int[2])
```

# Plot fold change results
```{r fig.width=3, fig.height=3}
ggplot(mean_ctcf, aes(x = strain, y = `avgCTCF_ELT-2`/avgCTCF_L4440)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  labs(title = "elt-2 Promoter\nRelative Fluorescence") +
  theme_classic()+
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.5),
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 16),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size=20),
    plot.title = element_text(hjust = 0.5, size = 20)
    ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  ylab("Relative Fluorescence") + coord_flip()

ggsave("../03_output/ELT-2_promoter_analysis_fold_change_210317.pdf", width = 3, height = 3, useDingbats=FALSE)  
```
